---
title: "Thanos Snap Template"
output: word_document
names: "Andrew S, Lucas S, Noah C, Dillon K"
date: "2023-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, include=FALSE}
## Import libraries
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(dslabs)
library(gridExtra)
library(ggrepel)
library(ggthemes)
library(knitr)
library(viridis)
library(stringr)
library(rvest)
library(hexbin)
library(ggplot2)
library(gganimate)
library(gifski)
library(png)
library(ggforce)
```


## Introduction

Our project task was to learn about genetic drift and the affects of different population aspects on genetic drift. We then needed to simulate this with our own population and summarize the results in an interactive Shiny document. The following functions were created to accomplish the population data simulation. This information was then used in the Shiny document we made with some modifications to finish up the outputs of the plots and table along with the smooth running of the simulation.


## Function 1

```{r, eval=FALSE}
# Function to create a generation of offspring, or initial population
# Inputs: generation size 'size', alleles in genetic pool 'alleles', and allele proportion 'allele.prop'
# Outputs: tibble with a generation of offspring

create.generation <- 
  function(size=100, alleles=c("AA", "Aa", "aa"), allele.prop=c(.25,.5,.25)) {
    gen <- tibble(Trait = NA, Allele = NA, .rows = size)
    gen$Allele <- sample(alleles, size = size, prob = allele.prop, replace = TRUE)
    for(index in 1:nrow(gen)) {
      if(gen$Allele[index] == "AA" | gen$Allele[index] == "Aa") {
        gen$Trait[index] <- "Dominant"
      } else {
        gen$Trait[index] <- "Recessive"
      }
    }
    return(gen)
  }
```

Function 1:   'create.generation' is a function to create a generation of observations, each having it's own Genotype and Trait. The function creates the generation based on the size that's desired, genotypes provided, and proportion that each genotype has in the population. Then a random sample of the genotypes, labeled 'alleles' in the function, is taken with the given probabilities of each one. Finally, the Trait, either Dominant or Recessive, is filled in to match the genotypes.


## Function 2

```{r, eval=FALSE}
# Function to induce bottleneck event and alter population
# Inputs: tibble with population to affect 'popu', and type of bottleneck event 'event', survival rate input 'survival.rate' changes survival rate of pathogen for each characteristic and how drastic natural distasters are
# Outputs: the tibble with the resulting population (the survivors of the population)

bottleneck <- function(popu, event, survival.rate) {
  if(event == "Founder" ) {
    survival.rate <- NA
    deaths <- sample(1:nrow(popu), size = nrow(popu)*.9)
    
  } else if(event == "Dominant Characteristic Pathogen") {
    dom.index <- which(popu$Trait=="Dominant")
    deaths <- sample(dom.index, size = survival.rate*length(dom.index))
    
  } else if(event == "Recessive Characteristic Pathogen") {
    rec.index <- which(popu$Trait=="Recessive")
    deaths <- sample(rec.index, size = survival.rate*length(rec.index))
    
  } else if(event == "Natural Disaster") {
    deaths <- sample(1:nrow(popu), size = survival.rate*nrow(popu))
    
  } else if(event == "Thanos Snap") {
    deaths <- sample(1:nrow(popu), size = .5*nrow(popu))
  }
  
  affected <- popu[-deaths,]
  return(affected)
}
```

Function 2:   'bottleneck' is a function that takes a population and induces a bottleneck event on it. Each bottleneck event has a different way of killing off portions of the population. For each option a random sample of people is taken from the population and killed. Then the population is filtered so only the survivors are left in the table.


## Function 3

```{r, eval=FALSE}
# Function to get alleles and allele proportions from population
# Inputs: tibble with alleles and proportions 'pop'
# Outputs: tibble with the alleles and their respective proportions from population

get.proportions <- function(pop) {
  proportions <- pop %>% group_by(Allele) %>% summarize(prop = length(Allele)/nrow(pop))
  return(proportions)
}
```

Function 3:   'get.proportions' is a function to get the genotype frequencies out of a population or generation. The function takes a population, groups by their genotype, "Allele", and summarizes the information so the proportions are acquired.


## Function 4

```{r, eval=FALSE}
## Sample simulation
# Function to simulate the generations
# Inputs: number of generations 'gens', default is 10, the starting allele frequency 'start.freq' ordered respective to 'AA', 'Aa', 'aa'. Also population size 'pop.size' and bottleneck event 'bottle' and survival rate for an event, specific to characteristic if either pathogen is chosen 'surv.rate'
# Outputs: the resulting final generation of the simulation in a tibble

simulate.generations <- function(gens=10, start.freq=c(.25, .5, .25), pop.size=100, bottle="Founder", surv.rate = NA) {
  population <- create.generation(size = pop.size, allele.prop = start.freq)
  all.gens <- population %>% mutate(Gen = 1)
  for(gen in 2:gens) {
    props <- population %>% get.proportions()
   
    # Alteration
     if(gen == gens/2 | gen == (gens-1)/2) {
      population <- create.generation(size = pop.size, alleles=props$Allele, allele.prop=props$prop) %>% 
        bottleneck(event = bottle, survival.rate = surv.rate)
    } else {
    
     if(nrow(population) < pop.size) {
       population <- create.generation(size = (nrow(population)+0.10*pop.size), alleles=props$Allele, allele.prop=props$prop)
     } else {
       population <- create.generation(size = pop.size, alleles=props$Allele, allele.prop=props$prop)
     }
    }
    population <- population %>% mutate(Gen = gen)
    all.gens <- all.gens %>% rbind(population)
  }
  return(all.gens)
}
```

Function 4:   'simulate.generations' is the function that the simulation runs on and includes all the previous functions. It starts with a population (the first generation) and gets the proportions of genotypes to create a succeeding generations that replaces the generation that the proportions were taken from. Each generation is also added to another table that holds all the generations of the simulation. Halfway through the desired number of generations, the bottleneck event is induced on the population and the remaining generations are the result of the bottleneck.

## Shiny Code

```{r, eval=FALSE}
# Define UI ----

# Shiny ideas from https://github.com/santiagosnchez/genetic_drift
ui <- fluidPage(
    titlePanel("Genetic Drift Simulator"),
    sidebarLayout(
        sidebarPanel(
            numericInput(
                inputId = "population_size", 
                label = "Population Size", 
                value = 100, 
                step = 1, 
                min = 2,
                max = 1000
            ), 
            numericInput(
                inputId = "num_generations",
                label = "Number of Generations",
                value = 10,
                step = 1,
                min = 1,
                max = 50
            ),
            selectInput(
                inputId = "allele",
                label = "Allele to Manipulate",
                choices = c("A", "a"),
                selected = "A",
            ),
            sliderInput(
                inputId = "initial_frequency", 
                label = "Initial Allele Frequency", 
                min = 0, 
                max = 1, 
                step = 0.05, 
                value = 0.50
            ), 
            selectInput(
                inputId = "bottleneck",
                label = "Bottleneck Event",
                choices = c("Thanos Snap", "Founder", "Dominant Characteristic Pathogen", "Recessive Characteristic Pathogen", "Natural Disaster"),
                selected = "Thanos Snap"
            ),
            sliderInput(
                inputId = "mortality_rate",
                label = "Mortality Rate for Natural Disasters and Pathogens",
                min = 0.05,
                max = 1,
                step = 0.05,
                value = 0.5,
            ),
            actionButton(
                inputId = "run", 
                label = "Run"), 
            actionButton(
                inputId = "reset", 
                label = "Reset"), 
            ),
        mainPanel(
            plotOutput(
                outputId = "plots"
            ),
            tableOutput(
                outputId <- "table"
            ),
            textOutput(
                outputId = "bottle.description"
            )
        )
    )
)

# Define server logic ----
server <- function(input, output, session) {
    library(tidyverse)
    summary.table <- tibble(Genotype = c("AA", "Aa", "aa"))
  
    observeEvent(input$run, {
        # Start simulation
        source("FinalFunctions.R")
  
        # Generating Starting Allele Frequencies
        if(input$allele == "A") {
            A.freq <- input$initial_frequency
            a.freq <- 1 - A.freq
        } else {
            a.freq <- input$initial_frequency
            A.freq <- 1 - a.freq
        }
    
        # Running main simulation code
        population <- simulate.generations(
        gens = input$num_generations, 
        start.freq = c(A.freq^2, 2*A.freq*a.freq, a.freq^2), 
        pop.size = input$population_size, 
        bottle = input$bottleneck, 
        surv.rate = input$mortality_rate)
  
        # Cleaning up data for plots
        plotsInfo <- population %>% group_by(Gen, Allele) %>% summarize(n.Allele = length(Allele))
        genSizes <- population %>% group_by(Gen) %>% summarize(n.Gen = length(Gen))
        plotsInfo <- plotsInfo %>% left_join(genSizes, by = "Gen") %>% mutate(proportions = n.Allele/n.Gen)
        
        # Generated plots
        scatter <- plotsInfo %>%
        ggplot(aes(Gen, n.Allele)) + 
        geom_point(aes(color = Allele), size = 3) + 
        labs(x = "Generation", y = "Genotype Count", title = "Genetic Drift Simulation", color = "Genotype", size = "")
  
  
        # Modify summary table
        prop.change <- plotsInfo %>% filter(Gen == 1 | Gen == input$num_generations) %>% 
        group_by(Allele) %>% 
        summarize("Proportion Change" = proportions[which(Gen == input$num_generations)] - proportions[which(Gen == 1)])
        colnames(prop.change) <- c("Genotype", "Proportion Change")
        summary.table <- summary.table %>% left_join(prop.change, by = "Genotype")
    
        # Output summary plot
        output$plots <- renderPlot({
            scatter
        })
  
        # Generate information to display about bottleneck event
        ifelse(input$bottleneck == "Thanos Snap", 
            description <- "Boom! Thanos snapped his fingers and half the population has been wiped out!",
            ifelse(input$bottleneck == "Founder", 
                description <- "One tenth of the population isolated itself and repopulated.",
                ifelse(input$bottleneck == "Dominant Characteristic Pathogen",
                    description <- "A Deadly Pathogen that only affects those with the dominant characteristic is introduced 
                                    into the population! Only those with genotype AA or Aa affected.",
                    ifelse(input$bottleneck == "Recessive Characteristic Pathogen",
                        description <- "A Deadly Pathogen that only affects those with the recessive characteristic is introduced 
                                        into the population! Only those with genotype aa affected.", 
                        description <- "A Natural Disaster struck the area that the population resides and wiped out a portion of the population!"
                    )
                )
            )
        )

        # Display bottleneck description
        output$bottle.description <- renderText({
            description
        })
    
        # Output updated Table
        output$table <- renderTable({
            summary.table
        })

    })

    # Reset simulation
    observeEvent(input$reset, {
        # Stop simulation and clear progress
        population <- NULL
        plotsInfo <- NULL
        genSizes <- NULL
        animation <- NULL
        output$plots <- renderPlot({})
        output$bottle.description <- renderText({})
        output$table <- renderTable({})
    })

    # Output summary table
    output$table <- renderTable({
        summary.table
    })
}

# Run the app ----
shinyApp(ui = ui, server = server)

```

Shiny code:   The shiny document made it possible for us to make an interactive HTML page where users can change the population size, number of generations, starting allele proportions, bottleneck event type, and mortality rate for certain bottleneck events. Our app consists of the User Interace and server logic. Our HTML outputs the summary plot with Genotype count on the Y axis and Generation number on the X axis with the Genotypes identified by color. The summary table shows the difference in population proportion of Genotypes over the span of the simulation. The last generation proportions minus the starting generation proportions.


## Conclusion

This project was interesting and offered various challenges. There is definitely more information we could incorporate into the final Shiny document and different ways we might be able to show Genetic Drift more accurately. Overall, the Shiny document turned out well and learning to use Shiny will likely be an influential skill to have. We would recommend this project to future groups. 
