---
title: "Thanos Snap Template"
output: word_document
date: "2023-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, include=FALSE}
## Import libraries
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(dslabs)
library(gridExtra)
library(ggrepel)
library(ggthemes)
library(knitr)
library(viridis)
library(stringr)
library(rvest)
library(hexbin)
library(ggplot2)
library(gganimate)
library(gifski)
library(png)
library(ggforce)
```


## Introduction

Intro text here



## Function 1

```{r, eval=FALSE}
# Function to create a generation of offspring, or initial population
# Inputs: generation size 'size', alleles in genetic pool 'alleles', and allele proportion 'allele.prop'
# Outputs: tibble with a generation of offspring

create.generation <- 
  function(size=100, alleles=c("AA", "Aa", "aa"), allele.prop=c(.25,.5,.25)) {
    gen <- tibble(Trait = NA, Allele = NA, .rows = size)
    gen$Allele <- sample(alleles, size = size, prob = allele.prop, replace = TRUE)
    for(index in 1:nrow(gen)) {
      if(gen$Allele[index] == "AA" | gen$Allele[index] == "Aa") {
        gen$Trait[index] <- "Dominant"
      } else {
        gen$Trait[index] <- "Recessive"
      }
    }
    return(gen)
  }
```

Function 1:   'create.generation' is a function to create a generation of observations, each having it's own Genotype and Trait. The function creates the generation based on the size that's desired, genotypes provided, and proportion that each genotype has in the population. Then a random sample of the genotypes, labeled 'alleles' in the function, is taken with the given probabilities of each one. Finally, the Trait, either Dominant or Recessive, is filled in to match the genotypes.


## Function 2

```{r, eval=FALSE}
# Function to induce bottleneck event and alter population
# Inputs: tibble with population to affect 'popu', and type of bottleneck event 'event', survival rate input 'survival.rate' changes survival rate of pathogen for each characteristic and how drastic natural distasters are
# Outputs: the tibble with the resulting population (the survivors of the population)

bottleneck <- function(popu, event, survival.rate) {
  if(event == "Founder" ) {
    survival.rate <- NA
    deaths <- sample(1:nrow(popu), size = nrow(popu)*.9)
    
  } else if(event == "Dominant Characteristic Pathogen") {
    dom.index <- which(popu$Trait=="Dominant")
    deaths <- sample(dom.index, size = survival.rate*length(dom.index))
    
  } else if(event == "Recessive Characteristic Pathogen") {
    rec.index <- which(popu$Trait=="Recessive")
    deaths <- sample(rec.index, size = survival.rate*length(rec.index))
    
  } else if(event == "Natural Disaster") {
    deaths <- sample(1:nrow(popu), size = survival.rate*nrow(popu))
    
  } else if(event == "Thanos Snap") {
    deaths <- sample(1:nrow(popu), size = .5*nrow(popu))
  }
  
  affected <- popu[-deaths,]
  return(affected)
}
```

Function 2:   'bottleneck' is a function that takes a population and induces a bottleneck event on it. Each bottleneck event has a different way of killing off portions of the population. For each option a random sample of people is taken from the population and killed. Then the population is filtered so only the survivors are left in the table.


## Function 3

```{r, eval=FALSE}
# Function to get alleles and allele proportions from population
# Inputs: tibble with alleles and proportions 'pop'
# Outputs: tibble with the alleles and their respective proportions from population

get.proportions <- function(pop) {
  proportions <- pop %>% group_by(Allele) %>% summarize(prop = length(Allele)/nrow(pop))
  return(proportions)
}
```

Function 3:   'get.proportions' is a function to get the genotype frequencies out of a population or generation. The function takes a population, groups by their genotype, "Allele", and summarizes the information so the proportions are acquired.


## Function 4

```{r, eval=FALSE}
## Sample simulation
# Function to simulate the generations
# Inputs: number of generations 'gens', default is 10, the starting allele frequency 'start.freq' ordered respective to 'AA', 'Aa', 'aa'. Also population size 'pop.size' and bottleneck event 'bottle' and survival rate for an event, specific to characteristic if either pathogen is chosen 'surv.rate'
# Outputs: the resulting final generation of the simulation in a tibble

simulate.generations <- function(gens=10, start.freq=c(.25, .5, .25), pop.size=100, bottle="Founder", surv.rate = NA) {
  population <- create.generation(size = pop.size, allele.prop = start.freq)
 # population <- population %>% bottleneck(event = bottle, survival.rate = surv.rate)
  all.gens <- population %>% mutate(Gen = 1)
  for(gen in 2:gens) {
    props <- population %>% get.proportions()
   
    # Alteration
     if(gen == gens/2) {
      population <- create.generation(size = pop.size, alleles=props$Allele, allele.prop=props$prop) %>% 
        bottleneck(event = bottle, survival.rate = surv.rate)
    } else {
    
     if(nrow(population) < pop.size) {
       population <- create.generation(size = (nrow(population)+0.10*pop.size), alleles=props$Allele, allele.prop=props$prop)
     } else {
       population <- create.generation(size = pop.size, alleles=props$Allele, allele.prop=props$prop)
     }
    }
    population <- population %>% mutate(Gen = gen)
    all.gens <- all.gens %>% rbind(population)
  }
  return(all.gens)
}
```

Function 4:   'simulate.generations' is the function that the simulation runs on and includes all the previous functions. It starts with a population (the first generation) and gets the proportions of genotypes to create a succeeding generations that replaces the generation that the proportions were taken from. Each generation is also added to another table that holds all the generations of the simulation. Halfway through the desired number of generations, the bottleneck event is induced on the population and the remainding generations are the result of the bottleneck.


## Conclusion

Conclusion text here.









