---
title: "Thanos Snap Template"
output: word_document
date: "2023-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Two of us handle the R side of things to develop and create simulations. Two of us learn how to make shiny apps

2. Investigate different add-ons to our simulation such as population size, allele frequency, generations, events, etc. and how they impact genetic drift.

3. Continue to learn shiny. Start to create test shiny apps.

4. Create simulations in R for genetic drift after a bottleneck event. Then translate that simulation into something user-friendly.

5. Finished product: Shiny app that simulates genetic drift after a bottleneck event.


```{r}
## Import libraries
source("https://github.com/mennodejong1986/SambaR/raw/master/SAMBAR_v1.09.txt")
```



## Section One

```{r Andrew}
# Comment in code chunk
```
is annoying
is great
## Section Two

```{r Lucas}
ayo  = c(69, 69)
```
is weak
## Section Three

```{r Noah}
v1 <- c(6,9,1)
v1
```
is dope
is a girl

## Section Four

```{r Dillon}

```
Vikings suck