---
title: "Thanos Snap Template"
output: word_document
date: "2023-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



1. Two of us handle the R side of things to develop and create simulations. Two of us learn how to make shiny apps

2. Investigate different add-ons to our simulation such as population size, allele frequency, generations, events, etc. and how they impact genetic drift.

3. Continue to learn shiny. Start to create test shiny apps.

4. Create simulations in R for genetic drift after a bottleneck event. Then translate that simulation into something user-friendly.

5. Finished product: Shiny app that simulates genetic drift after a bottleneck event.


```{r, warning=FALSE}
## Import libraries
library(tidyverse)
library(dplyr)
library(shiny)
library(ggplot2)
```



## Section One

```{r Andrew}
## One Generation of Survivors

# Total population for simulation
pop <- tibble(Trait = NA, Allele = NA, .rows = 100) #set .rows to 'population_size'

pop$Allele <- sample(c("AA", "Aa", "aa"), size=nrow(pop), replace=TRUE, prob=c(0.25, 0.5, 0.25)) #Eventually prob will be set by initial_frequency

# Random index for survival selection
survivors1 <- sample(1:nrow(pop), size = nrow(pop)/2)

# Keeping only the survivors
pop <- pop[survivors1,]

## Adding Generations

prop <- pop %>% group_by(Allele) %>% summarize(proportion = length(Allele)/nrow(pop))

```

```{r generation}
# Function to create a generation of offspring, or initial population
# Inputs: generation size 'size', alleles in genetic pool 'alleles', and allele proportion 'allele.prop'
# Outputs: tibble with a generation of offspring
create.generation <- 
  function(size=100, alleles=c("AA", "Aa", "aa"), allele.prop=c(.25,.5,.25)) {
  gen <- tibble(Trait = NA, Allele = NA, .rows = size)
  gen$Allele <- sample(alleles, size = size, prob = allele.prop, replace = TRUE)
    for(index in 1:nrow(gen)) {
     if(gen$Allele[index] == "AA" | gen$Allele[index] == "Aa") {
       gen$Trait[index] <- "Dominant"
     } else {
       gen$Trait[index] <- "Recessive"
     }
   }
  return(gen)
  }
```

```{r}
# Test Function 'create.generation'
generation <- create.generation()

# Testing function still
initial.pop <- create.generation()

# Getting info from population
pop.info <- get.proportions(initial.pop)

# Creating another generation 'gen2' using population info
nextGen <- create.generation(size = nrow(initial.pop), alleles=pop.info$Allele, allele.prop = pop.info$prop)

# Possible binding of initial population with second generation
initial.pop <- initial.pop %>% rbind(nextGen)

## Replacing initial population with next generation
# initial.pop <- nextGen

```

```{r bottleneck}
# Function to induce bottleneck event and alter population
# Inputs: tibble with population to affect 'popu', and type of bottleneck event 'event', survival rate input 'survival.rate' changes survival rate of pathogen for each characteristic and how drastic natural distasters are
# Outputs: the tibble with the resulting population (the survivors of the population)
bottleneck <- function(popu, event, survival.rate) {
  if(event == "Founder" ) {
    survival.rate <- NA
    deaths <- sample(1:nrow(popu), size = nrow(popu)*.9)
    
  } else if(event == "Dominant Characteristic Pathogen") {
    dom.index <- which(popu$Trait=="Dominant")
    deaths <- sample(dom.index, size = survival.rate*length(dom.index))
    
  } else if(event == "Recessive Characteristic Pathogen") {
    rec.index <- which(popu$Trait=="Recessive")
    deaths <- sample(rec.index, size = survival.rate*length(rec.index))
    
  } else if(event == "Natural Disaster") {
    deaths <- sample(1:nrow(popu), size = survival.rate*nrow(popu))
    
  } else if(event == "Thanos Snap") {
    deaths <- sample(1:nrow(popu), size = .5*nrow(popu))
  }
  
  affected <- popu[-deaths,]
  return(affected)
}
```

```{r pop.info}
# Function to get alleles and allele proportions from population
# Inputs: tibble with alleles and proportions 'pop'
# Outputs: tibble with the alleles and their respective proportions from population
get.proportions <- function(pop) {
  proportions <- pop %>% group_by(Allele) %>% summarize(prop = length(Allele)/nrow(pop))
  return(proportions)
}
```

```{r simulation}
## Sample simulation
# Function to simulate the generations
# Inputs: number of generations 'gens', default is 10, the starting allele frequency 'start.freq' ordered respective to 'AA', 'Aa', 'aa'. Also population size 'pop.size' and bottleneck event 'bottle' and survival rate for an event, specific to characteristic if either pathogen is chosen 'surv.rate'
# Outputs: the resulting final generation of the simulation in a tibble
simulate.generations <- function(gens=10, start.freq=c(.25, .5, .25), pop.size=100, bottle="Founder", surv.rate = NA) {
  population <- create.generation(size = pop.size, allele.prop = start.freq)
  population <- population %>% bottleneck(event = bottle, survival.rate = surv.rate)
  all.gens <- population %>% mutate(Gen = 1)
  for(gen in 2:gens) {
    props <- population %>% get.proportions()
    if(nrow(population) < pop.size) {
      population <- create.generation(size = (nrow(population)+0.5*pop.size), alleles=props$Allele, allele.prop=props$prop)
    } else {
       population <- create.generation(size = pop.size, alleles=props$Allele, allele.prop=props$prop)
    }
    population <- population %>% mutate(Gen = gen)
    all.gens <- all.gens %>% rbind(population)
  }
  return(all.gens)
}
```

```{r testing}
## Testing Simulation

# Simulation using Founder Effect
Founder.sim <- simulate.generations()

# Simulation using Dominant Characteristic Pathogen
dom.path.sim <- simulate.generations(bottle = "Dominant Characteristic Pathogen", surv.rate = 0.8)

# Simulation using Recessive Characteristic Pathogen
rec.path.sim <- simulate.generations(bottle = "Recessive Characteristic Pathogen", surv.rate = 0.8)

# Simulation using Natural Disaster
nat.disaster.sim <- simulate.generations(bottle = "Natural Disaster", surv.rate = 0.5)

# Simulation using Thanos Snap
than.snap.sim <- simulate.generations(bottle = "Thanos Snap")


```

```{r summarizing}

Founder.sim %>% group_by(Gen, Allele) %>% summarize(props = length(Allele)/nrow(Founder.sim)*10) %>% ggplot(aes(Gen, props)) + geom_col(aes(fill = Allele), position = "dodge")

dom.path.sim %>% group_by(Gen, Allele) %>% summarize(props = length(Allele)/nrow(dom.path.sim)*10) %>% ggplot(aes(Gen, props)) + geom_col(aes(fill = Allele), position = "dodge")

rec.path.sim %>% group_by(Gen, Allele) %>% summarize(props = length(Allele)/nrow(rec.path.sim)*10) %>% ggplot(aes(Gen, props)) + geom_col(aes(fill = Allele), position = "dodge")

```

## Section Two

```{r Lucas}

```

## Section Three

```{r Noah}

```

## Section Four

```{r Dillon}
# Currently working on shiny stuff.

# Attempt at a bottleneck function

run.bottleneck = function(prop.lost=0.50){ # Eventually will be controlled by a slider/text input
    # Random index for survival selection
    survivors <- sample(1:nrow(pop), size = nrow(pop) * (1-prop.lost))
    
    # Keeping only the survivors
    pop <- pop[survivors,]
    
    # Somehow continue simulation (maybe a resume button?)
}
```